---
title: "Class 5: Data Visualization with GG Plot"
author: "Candyd Lace R. Velasquez"
format: pdf
toc: true
---

# Our first plot

R has base graphics

```{r}
plot(cars)
```

How would I plot this with `ggplot2`? No, we need to install ggplot package first

To install any package use `install.packages()` function make sure to load up the package in the session you are using use `library()` function

```{r}
#install.packages("ggplot2")
```

Every ggplot needs at least 3 layers :

\-**Data**(ie the data.frame we have),

\-**Aes** (the aesthetic mapping of our data to what we want to plot)

\-**Geoms** (How we want to plot this stuff!)

```{r}
library(ggplot2)
ggplot(data=cars) +
  aes(x=speed, y=dist) +
  geom_point()
```

```{r}
library(ggplot2)
```

```{r}
ggplot(data=cars) +
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

### A more interesting plot

First read

Q1. how many genes are in this dataset

```{r}
nrow(genes)
```

```{r}
colnames(genes)
```


```{r}
ncol(genes)
```

```{r}
table(genes$State)
```

```{r}
round(table(genes$State)/nrow(genes) * 100, 3)
```


There are `r nrow(genes)` in this data set

```{r}
ggplot(data=genes) +
  aes(x=Condition1, y=Condition2, col=State)+ 
  geom_point()
```

```{r}
p <-ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State)+ 
  geom_point()
p
```

```{r}
p + scale_colour_manual( values=c("orchid","gray","green") )

```




```{r}
p + scale_colour_manual(values=c("orchid","gray","green")) +
    labs(title="Gene Expresion Changes Upon Drug Treatment",
         x="Control (no drug) ",
         y="Drug Treatment")
```