---
title: "Class 10: Halloween Mini-Project"
author: "Candyd Lace"
date: 10/28/2022
format: pdf


---

# Background:

In this mini-project we will examine Halloween Candy data. What is your favorite candy? What is nougat anyway? How do you say it in America?

FIrst step is to read the data

```{r}

candy <- read.csv("candy-data.txt", row.names=1)
head(candy)

```

# Q How many different candy types are in this data set ?

```{r}
nrow(candy)
```

#Q2 How many fruity candy types are in the data set

```{r}
sum(candy$fruity)
```

# win percent of your favorite candy

```{r}
rownames(candy)
```

```{r}
candy["Sugar Babies", ]$winpercent
```

```{r}
candy["WertherÕs Original Caramel", ]
```

```{r}
library("skimr")
skim(candy)
```

# Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

win percent

# Q7. What do you think a zero and one represent for the candy\$chocolate column?

# Q8. Plot a histogram of winpercent values:

```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)
ggplot(candy, aes(x=winpercent)) +geom_histogram(bins=10, col="blue", fill="violet")
```

\# Q9 is the distribution of winpercent value symmetrical?

no

\# Q10 is the center of the distribution above or below 50%?

below 50%

\# Q11 On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate.inds <-as.logical(candy$chocolate)
chocolate.win <-candy[chocolate.inds,]$winpercent

mean(chocolate.win)
```

# Q12 Is this statistically significant

Yes

```{r}

fruity.inds <-as.logical(candy$fruity)
fruity.win <-candy[fruity.inds,]$winpercent

mean(fruity.win)
```

```{r}
t.test(chocolate.win, fruity.win)
```

# 3. Overall Candy Rankings

The base R `sort()` and `order()` functions are very useful

```{r}
x <-c(5,1,2,6)

sort(x)
```

order function: tells you the position

```{r}
x[order(x)]
```

```{r}
y <-c("barry", "alice", "chandra")
y
sort(y)
order(y)
```

# Q13 What are the five least candy types in this set?

First I want to order/arrange the entire dataset by winpercent values

Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
inds <-order(candy$winpercent)
head(candy[inds,], n=5)
```

```{r}
head(candy[order(candy$winpercent),], n=5)
```

# Q15 make a first barplot of candy ranking based on winpercent values

Barplot: the dafult barplot, made with `geom_col()` has the bars in order

```{r}
p <-ggplot(candy) + aes(winpercent, reorder( rownames(candy), winpercent)) + geom_col()
```

```{r}
ggsave("mybarplot.png")
```

# Create color vector: color every single bar in the plot

```{r}
my_cols <- rep("black", nrow(candy))
#my_cols
my_cols[as.logical(candy$chocolate)] <-"chocolate"
my_cols[as.logical(candy$bar)] <-"brown"
my_cols[as.logical(candy$fruity)] <-"violet"
my_cols
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```


# Taking a look at pricepercent
Q19. Which candy type is the highest ranked in terms of winpercent for the least money
Reese's miniature
Q20: what are the top 5 msot expensive candy types in the dataset and of these which is least popular?

Nik L Nip
Nestle Smarties
Ring Pop
Sugar Babies
POprocks 

What about value for money?
What is the best candy for the least money?

One way would be to plot `winpercent` vs the `pricepercent`

```{r}
ggplot(candy)+
  aes(winpercent, pricepercent, label=rownames(candy)) +geom_point(col= my_cols) +
geom_text(col=my_cols)
```


This plot sucks. Can't read the labels 
Use ggrepl package to help 

```{r}
library(ggrepel)
ggplot(candy)+
  aes(winpercent, pricepercent, label=rownames(candy)) +geom_point(col= my_cols) +
geom_text_repel(col=my_cols, size=2.5, max.overlaps=7)
```

# 5 Exploring the correlation structure 

```{r}
library(corrplot)
```
```{r}
cij <-cor(candy)
corrplot(cij)
```

Q22: Examining this plot what two variables are anti-correlated (ie have minus values?)
chocolate and fruity 



Q23: Similarly, what two variables are most positively correlated?
chocolate and how popular it is or bar


# 6 PCA: Principal Component Analysis
The main function that always there for us is `prcomp()`. It has an important arguemtn that is selt to `scale=FALSE`


```{r}
pca <-prcomp(candy, scale=TRUE)
summary(pca)
```

my PCA plot (a.k.a) PC1 vs PC2 score plot 

```{r}
plot(pca$x[,1], pca$x[,2], col=my_cols, pch=16)
```

make a nicer plot with gg plot 
```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```


```{r}
library(plotly)
#ggplotly(p)
```


```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

# Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard and pluribus
